---
title: "Hexoskin"
author: "Jonathan Slaney"
date: "03/10/2019"
output: html_document
---

## Necessary packages

```{r}
library(knitr)
library(tidyverse)
library(tidymodels)
library(ggmap)
library(lubridate)
library(readr)
library(psych)
library(ggpubr)
library(ggpmisc)
library(effsize)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Looping over data and creating city and participant columns

```{r}
hxd <- read_csv("/Users/dfuller/Desktop/DatasetsHexoskin/hexo_data_analysis.csv")

### Descriptives (for heart rate, breathing) psych package describe function
```

```{r}
describe(hxd$hr)

ggplot(hxd, aes(hr)) + 
    geom_histogram()
```

```{r}
describe(hxd$ve_L)

ggplot(hxd, aes(ve_L)) + 
    geom_histogram()
```

### Descriptives for Participants and Cities

```{r}
table(hxd$participants)

table(hxd$city)

table(hxd$city, hxd$participants)
```

### Histograms by city 

```{r}
ggplot(hxd, aes(hr)) +
  geom_histogram(aes(y = ..density..), fill = "cornflowerblue", colour = "black", size = 0.4) +
  geom_density(aes(y = ..density..), colour = "darkred", size = 1) + 
  ylab("HR density") + 
  xlab("Heart rate (bpm)") + 
   ggtitle("HR by city") + 
   facet_wrap(~ city)
```

### Histograms by city for Ve

```{r}
ggplot(hxd, aes(ve_L)) +
  geom_histogram(aes(y = ..density..), fill = "cornflowerblue", colour = "black", size = 0.4) +
  geom_density(aes(y = ..density..), colour = "darkred", size = 1) + 
  ylab("Density") + 
  xlab("Ve (L/min)") + 
   ggtitle("Ve by city") + 
   facet_wrap(~ city)
```

### Histogram of all data

```{r}
ggplot(hxd, aes(hr)) +
  geom_histogram(aes(y = ..density..), fill = "cornflowerblue", colour = "black", size = 0.4) +
  geom_density(aes(y = ..density..), colour = "darkred", size = 1) + 
  ylab("HR density") + 
  xlab("Heart rate (bpm)")
```

### Linear regressions by participant

### scatterplot for predictor and outcome

```{r}
cor_bar <- ggplot(hxd, aes(x = hr, y = ve_L)) + 
                  geom_point(size = 0.05) + 
  geom_smooth(method='lm', formula = y ~ x) + 
   facet_wrap(~ city)
plot(cor_bar)
```

### Calculate R^2

### Fit linear regression to participants

```{r}
cor_hexo_all <- ggplot(hxd, aes(x = hr, y = ve_L)) + 
                  geom_point(size = 0.1) + 
   geom_smooth(method='lm', formula = y ~ x) + 
   facet_wrap(~ participants)
plot(cor_hexo_all)
```

Here VJ is strange. Appears to have two distinct lines. Might indicates that this is two participants mixed? 

### Fit linear regression to all data

### Simple linear regression for hr and Ve

```{r}
lm_hr_ve <- lm(ve_L ~ hr, data = hxd)
summary(lm_hr_ve)
fit_m1 <- glance(lm_hr_ve)
hxd$lm_fitted <- predict(lm_hr_ve)

ggplot(hxd, aes(x = hr, y = ve_L)) + 
   geom_point(size = 1, alpha = 0.05) + 
   geom_line(aes(y = lm_fitted), size = 1, colour = "red") +
   xlab("Cycling Intensity (bpm)") + 
   ylab("Breathing Volume (L/minute)")
```

### Simple linear regression for hr and Ve with VE squared

```{r}
lm_hr_ve2 <- lm(ve_L ~ hr + I(hr^2), data = hxd)
summary(lm_hr_ve2)
fit_m2 <- glance(lm_hr_ve2)
hxd$lm2_fitted <- predict(lm_hr_ve2)

ggplot(hxd, aes(x = hr, y = ve_L)) + 
   geom_point(size = 1, alpha = 0.05) + 
   geom_line(aes(y = lm2_fitted), size = 1, colour = "red") +
   xlab("Cycling Intensity (bpm)") + 
   ylab("Breathing Volume (L/minute)")
```

### Simple linear regression for hr and Ve with VE cubed

```{r}
lm_hr_ve3 <- lm(ve_L ~ hr + I(hr^2) + I(hr^3), data = hxd)
summary(lm_hr_ve3)
fit_m3 <- glance(lm_hr_ve3)
hxd$lm3_fitted <- predict(lm_hr_ve3)

ggplot(hxd, aes(x = hr, y = ve_L)) + 
   geom_point(size = 1, alpha = 0.05) + 
   geom_line(aes(y = lm3_fitted), size = 1, colour = "red") +
   xlab("Cycling Intensity (bpm)") + 
   ylab("Breathing Volume (L/minute)")
```

### Linear regression for hr and Ve with VE squared plus gender

```{r}
lm_hr_ve4 <- lm(ve_L ~ hr + I(hr^2) + sex, data = hxd)
summary(lm_hr_ve4)
fit_m4 <- glance(lm_hr_ve4)

hxd$lm4_fitted <- predict(lm_hr_ve4, newdata = hxd, level = 0)

ggplot(hxd, aes(x = hr, y = ve_L, color = sex) ) +
     geom_point(size = 1, alpha = 0.05) + 
     geom_line(data = hxd, aes(y = lm4_fitted), size = 1) +
      xlab("Cycling Intensity (bpm)") + 
      ylab("Breathing Volume (L/minute)")
```

### Linear regression for hr and Ve with VE squared plus gender plus city 

```{r}
lm_hr_ve5 <- lm(ve_L ~ hr + I(hr^2) + sex + city, data = hxd)
summary(lm_hr_ve5)
fit_m5 <- glance(lm_hr_ve5)

hxd$lm5_fitted <- predict(lm_hr_ve5, newdata = hxd, level = 0)

ggplot(hxd, aes(x = hr, y = ve_L, color = city) ) +
     geom_point(size = 1, alpha = 0.05) + 
     geom_line(data = hxd, aes(y = lm5_fitted), size = 1) +
      xlab("Cycling Intensity (bpm)") + 
      ylab("Breathing Volume (L/minute)")
```

### Getting fit statistics for each model 

```{r}
fitstats <- rbind(fit_m1, fit_m2, fit_m3, fit_m4, fit_m5)
fitstats
```


```{r}
city_regressions <- hxd %>% 
                     group_by(city) %>% 
                     do(tidy(lm(ve_L ~ hr + I(hr^2), .)))
city_regressions
```

## Regressions using MET outcome variable

### Simple linear regression for MET and Ve

```{r}
lm_hr_ve <- lm(ve_L ~ METs, data = MET)
summary(lm_hr_ve)
fit_mv1 <- glance(lm_hr_ve)
MET$lm_fitted <- predict(lm_hr_ve)

ggplot(MET, aes(x = METs, y = ve_L)) + 
   geom_violin() + 
   geom_line(aes(y = lm_fitted), size = 1, colour = "red") +
   xlab("Metabolic Equivelents of Task") + 
   ylab("Breathing Volume (L/minute)")
```

### Simple linear regression for MET and Ve with MET squared

```{r}
lm_hr_ve2 <- lm(ve_L ~ METs + I(METs^2), data = MET)
summary(lm_hr_ve2)
fit_mv2 <- glance(lm_hr_ve2)
MET$lm2_fitted <- predict(lm_hr_ve2)

ggplot(MET, aes(x = METs, y = ve_L)) + 
   geom_point(size = 1, alpha = 0.05) + 
   geom_line(aes(y = lm2_fitted), size = 1, colour = "red") +
   xlab("MET") + 
   ylab("Breathing Volume (L/minute)")
```

### Simple linear regression for MET and Ve with MET cubed

```{r}
lm_hr_ve3 <- lm(ve_L ~ METs + I(METs^2) + I(METs^3), data = MET)
summary(lm_hr_ve3)
fit_mv3 <- glance(lm_hr_ve3)
MET$lm3_fitted <- predict(lm_hr_ve3)

ggplot(MET, aes(x = METs, y = ve_L)) + 
   geom_point(size = 3, alpha = 0.1) + 
   geom_line(aes(y = lm3_fitted), size = 1, colour = "red") +
   xlab("MET") + 
   ylab("Breathing Volume (L/minute)")
```

### Simple linear regression for MET and Ve with MET cubed on a violin plot

```{r}
ggplot(MET, aes(x = METs, y = ve_L)) + 
   geom_violin(size = 3, alpha = 0.1) + 
   geom_line(aes(y = lm3_fitted), size = 1, colour = "red") +
   xlab("MET") + 
   ylab("Breathing Volume (L/minute)")
```

### Linear regression for MET and Ve with MET squared plus gender

```{r}
lm_hr_ve4 <- lm(ve_L ~ METs + I(METs^2) + sex, data = MET)
summary(lm_hr_ve4)
fit_mv4 <- glance(lm_hr_ve4)

MET$lm4_fitted <- predict(lm_hr_ve4, newdata = MET, level = 0)

ggplot(MET, aes(x = METs, y = ve_L, color = sex) ) +
     geom_point(size = 1, alpha = 0.05) + 
     geom_line(data = MET, aes(y = lm4_fitted), size = 1) +
      xlab("METs") + 
      ylab("Breathing Volume (L/minute)")
```

### Linear regression for MET and Ve with MET squared plus gender plus city 

```{r}
lm_hr_ve5 <- lm(ve_L ~ METs + I(METs^2) + sex + city, data = MET)
summary(lm_hr_ve5)
fit_mv5 <- glance(lm_hr_ve5)

MET$lm5_fitted <- predict(lm_hr_ve5, newdata = MET, level = 0)

ggplot(MET, aes(x = METs, y = ve_L, color = city)) +
     geom_point(size = 1, alpha = 0.05) + 
     geom_line(data = metric_vec_template(), aes(y = lm5_fitted), size = 1) +
      xlab("MET") + 
      ylab("Breathing Volume (L/minute)")
```

### Getting fit statistics for each model

```{r}
fitstats <- rbind(fit_mv1, fit_mv2, fit_mv3, fit_mv4, fit_mv5)
fitstats
```

```{r}
city_regressions <- MET %>% 
                     group_by(city) %>% 
                     do(tidy(lm(ve_L ~ METs + I(METs^2), .)))
city_regressions
```
